### 2016-03-17 paper.tex
Add introduction and genre definition
There are a few holes that need to be plugged in with values from TotAN, but that's going to be more easily done from the studio with the game on hand.

@@

Content goes here.
We might cite \citet{Koster2013}, but this example is really just a placeholder
so the bibtex portion of the build script succeeds.

>>>>

\newcommand{\totan}{\textit{TotAN}}
\newcommand{\smersh}{\textit{SMERSH}}

@@

\textit{Tales of the Arabian Nights}~\citep{Goldberg2009} is a 
tabletop board game based on the eponymous folk tale.
It was originally published in 1985~\citep{Goldberg1985} but is more commonly
known for its 2009 re-release. 
It is popular within the designer board game hobby with a board game
rank of 239 (thematic rank 75) on Board Game Geek as of this writing,
and Shut Up \& Sit Down rank it as the ninth best game of all 
time~\citep{ShutUp2015}.
In this game, players control a character within a mythical Arabian
setting, and players explore the known world to accumulate Story and
Destiny points. The winner is the one who is able to meet their 
Story and Destiny point goal while having a successful encounter
in Baghdad---the City of Peace.

An intriguing property of \textit{Tales of the Arabian Nights}~(hereafter, \textit{TotAN}) is that it eschews the conventional wisdom for game design,
that player immersion is related to agency.
When a player has an encounter in \totan, they are told the name of the
encounter---such as ``angry merchant,''  ``powerful prince,'' or
``\textbf{put location example here}''---and based only on this information,
they must choose a reaction from one of the given tables; for example, \textbf{put the content of one table here}.
Another player then reads a corresponding entry from 
the \textit{Book of Tales}, which is a \textbf{N}-page
tome containing \textbf{N} numbered entries.
Most entries provide introductory text followed by paragraphs
tagged with skills that the active player may choose to use---without
knowing the consequences. The reader then narrates the conclusion of the
encounter and informs the player of their rewards, often in terms of
Story or Destiny points, gaining new skills, finding treasures, or
gaining status cards that modify players' options in future encounters.
To summarize, the player chooses a location on the map, but they do not 
choose what is encountered; they choose a reaction to the encounter name
without knowing how it will be interpreted in the story;
if they have the corresponding skill, they may choose to use it but
without knowing whether it will help or hinder them.
\totan{} is not a game of skill, where the player with the best tactics
wins the game: it is somewhat arbitrary and imbalanced, but from its
design emerges a fanciful gameplay experience. It is not played to win
but rather to enjoy the emerging story with friends.

This property of \totan{} reflects several characteristics
of its source material,
as our team discovered from reading youth-friendly translations suggested in the
rulebook (specifically \citet{Neil1994}).
Like the original stories, \totan{} is a collection of shorter stories
that are sometimes linked together and sometimes not.
The characters in the folk tales are often victims of fate with
very little agency over their own encounters. The games rules follow
cultural norms expressed in the folk tales:
you cannot win while sex-changed or on pilgrimage, and both marriage
and children are great blessings---unless you have an ugly baby,
which is shameful.
Baghdad is the most important city in the world, where your adventure starts
and ends.
Furthermore, many stories in \totan{} come directly from the tales,
such as the Sindbad's escape from the valley of diamonds or Aladdin's
being trapped in a magic cave.

Playing \totan{}  inspired members of our team to read translations
of the original tales, which made us recognize the cleverness of the design
and its elegant dovetailing with the source material.
Furthermore, in reading the texts, we recognized that we had already
learned elements of the stories and the culture through playing the game,
although not always consciously or explicitly.
This inspired the following analysis, in which we tease apart the
various elements of \totan{}, considering them from both game design
and learning design points of view, and then share the results of a pilot
project to create a technology-enhanced game within the same genre.

\section{Defining the genre of player-narration games}

We propose that \totan{} is one of very few games comprising a genre
of \textit{player-narration games}, and we define this genre as having
the following characteristics.

\begin{enumerate}
\item \textit{Take place in a believable, consistent setting.} 
 Although the world of \totan{} may be unfamiliar to the player,
 everything in it is representative of the world described in the original
 folk tale---complete with the contradictions that add depth and nuance.

\item \textit{Use narrative as a primary feedback mechanism.} 
 In \totan{}, the player's reaction and skill choice yields two forms
 of feedback: the narrative description followed by the changes to the
 game state. The narrative is primary, both chronologically and aesthetically.
 Note that we follow \citet{Koster2012} in treating narrative as a 
 feedback mechanism, not a game mechanism.

\item \textit{Have measurable goals.} There is a winning condition as part
 of the social contract of play, following colloquially-accepted 
 standards for board games. The ``game'' is not simply constrained
 cooperative storytelling or a role-playing experience without formal end
 conditions.

\item \textit{Incorporate endogenously meaningful ambiguous decisions.}
 Following \citet{Burgun2012}, the decisions that one makes in the game
 are meaningful even though they are made without complete knowledge
 of the game state. For example, a player makes 
 the choice of reaction in \totan{} in hopes that it leads toward skills
 that they have, such as choosing ``Fight'' while in possession of the
 ``Weapon Use'' skill, even though one does not know whether or not
 this skill will have relevance in the resulting entry from the
 \textit{Book of Tales}.

\item \textit{Reward players for decision-making that reflects
    cultural understanding.}  The knowledge that a player brings to
  bear on decisions is not just knowledge of in-game systems.  If one
  chooses to Fight an Angry Ifrit in \totan{} without the Weapon Use
  skill, it is probably not going to end well. This prediction is not
  based on my having memorized the \textit{Book of Tales}, nor knowing how many
  hit dice an Ifrit has (as in \textit{Dungeons \& Dragons} powergaming).
  Rather, this prediction draws upon a cultural understanding of anger,
  spirits, violence, and likely much more.
  However, the game does this without establishing
  ``right'' and ``wrong'' (``moral/immoral'',
  ``light side/dark side'', ``paragon/renegade'') choices.  There
  remain unexpected twists: perhaps fighting the Angry Ifrit without
  combat skills makes him respect you and grant a boon.  \totan{},
  after all, is a game that allows the player to choose to ``drink'' a
  violent storm or ``enter'' a mysterious artifact.  While the world
  remains consistent, it also embraces the potential for surprising
  results.

\end{enumerate}

Two formal elements from \totan{} are intentionally not included in the
genre definition above. Perhaps the most clear omission is players' reading
each others' stories. While we agree that this is a critical element of 
\totan{}---and that it dovetails performatively 
and practomimetically~\citep{Travis2011} with Scheherazade's
performances in the source material---it is not clear that this is
characteristic to the genre.  Anecdotal evidence shows that
players with low reading skills can ruin the play experience for other
players, while those players could have read their own stories, aloud or not.
Similarly, \totan{} permits a single-player experience which, while
having a different aesthetic, still seems to produce the same general outcomes
we described above.

The other formal element omitted from the genre definition is the use of an
explicit map. The map is an important tangible aspect of \totan{}, 
in particular with its representation of Baghdad as the largest, most
important city in the world. However, we also note that virtual spaces
can be represented without such an explicit map, such as in the MUD family
of games or choose-your-own-adventure books, both of which
use discrete spaces without giving them visual manifestation.

Analyzing \textit{Agents of SMERSH}~\citep{Maxwell2012} establishes that
these characteristics form a genre and not merely selected elements from one
game.
\smersh{} is a cooperative game set in the Cold War, with players as 
covert agents trying to stop the enigmatic Dr.\ Lobo. 
Players move across the world collecting resources and battling henchmen,
facing encounters clearly inspired by \totan{}: flip a card to read an
encounter title, choose a reaction from a table, hear an introduction,
and then get different results based on skills. Unlike \totan{},
\smersh{} uses dice for story resolution, and supporting resource
management mechanisms introduce important strategic decisions.
Despite the use of dice for story resolution, the feedback is still
primarily narrative.

For contrast, consider \textit{Above and Below}~\citep{Laukat2015}, a
worker-management game about building a city both above and below the
surface.  \textit{Above and Below} is primarily about building an
economic engine that generates victory points faster than your
opponents within a fixed number of rounds. Exploring the underground
is one source of points, and this is done by determining a random
entry within its encounter book.  A player reads this entry aloud,
after which the player is presented with a decision of how to
react. However, the reaction titles are a thin wrapper around a
risk-and-reward balance: each has a number which is the number of
successes required on dice rolls, so players choose the highest number
they believe they can roll. Higher numbers always equate to better results,
and this roll is usually the end of the encounter: a successful player
is given an in-game reward with no other story or narrative.
Hence, although \textit{Above and Below} includes an encounter book
that players read to each other, it does not satisfy the characteristics
to be in the player-narration game genre.
